\documentclass[11pt,leqno,oneside]{amsbook}
\usepackage{tikz}
\usetikzlibrary{cd}
\usepackage{bbm}
\usepackage{todonotes}

\usepackage{./notes}
\usepackage{../ReAdTeX/readtex-core}
\usepackage{../ReAdTeX/readtex-dangerous}
\usepackage{../ReAdTeX/readtex-abstract-algebra}

\numberwithin{thm}{section}

\title[Artin-Wedderburn Theory]{Artin-Wedderburn Theory \\ Notes
  inspired by a class taught by Brian Parshall in Fall 2017}
\author{George H. Seelinger}
\date{Fall 2017}
\begin{document}
\maketitle
\section{Krull-Schmidt}
We start with an underated lemma.
\begin{thm}[Fitting's Lemma]
  Let \(0 \neq M\) be a left \(R\)-module with a composition series
  and \(f \from M \to M\) a module homomorphism. Since \(M\) has a
  composition series, there exists an
  \(n\) such that \[
    \im f \supset \im f^2 \supset \cdots \supset \im f^n = \im f^{n+1}
    = \cdots
  \]
  and \[
    \ker f \subset \ker f^2 \subset \cdots \subset \ker f^n = \ker
    f^{n+1} = \cdots 
  \]
  Define \(f^\infty(M) := \im f^n\) and \(f^{-\infty}(M) := \ker
  f^n\). Then, \[
    M \isom f^\infty(M) \oplus f^{-\infty}(M)
  \]
\end{thm}
\begin{proof}
  Let \(x \in f^\infty(M) \intersect f^{-\infty}(M)\). Then, \(x =
  f^n(y) = f^{2n}(z)\) and \(f^n(x) = 0\). Then, \(0 = f^n(x) =
  f^{2n}(y)\). Thus, \(y \in \ker f^{2n} = \ker f^n\), so \(x = f^n(y)
  = 0\). \\

  Now, let \(x \in M\). Then, \(x = [x-f^n(y)]+f^n(y)\) where \(f^n(y)
  \in f^\infty(M)\) and \(y\) such that \(f^{2n}(y) =
  f^n(x)\). Then, \[
    f^n(x-f^n(y)) = f^n(x) - f^{2n}(y) = f^n(x) - f^n(x) = 0
  \]
  Thus, \(x - f^n(y) \in f^{-\infty}(M)\) and \(f^n(y) \in
  f^\infty(M)\). 
\end{proof}
Now, we seek to prove the Krull-Schmidt theorem.
\begin{thm}[Krull-Schmidt Theorem]
  Let \(0 \neq M\) be an \(R\)-module which is both Artinian and
  Noetherian. Suppose
  \begin{align*}
    M & = M_1 \oplus \cdots \oplus M_h \\
    & = N_1 \oplus \cdots \oplus N_k
  \end{align*}
  where \(M_i, N_j\) are indecomposable. Then, \(h=k\) and, up to
  rearranging terms, \(M_i \isom N_i\) for \(i = 1, \ldots, h\). 
\end{thm}
To prove this theorem, we need the following lemmas.
\begin{lem}
  Let \(M,N\) be \(R\)-modules with \(N\) decomposable and
  homomorphisms \(\nu \from M \to N\) and \(\mu \from N \to M\). If
  \(\mu \nu \from M \to M\) is an automorphism, then \(\mu\) and
  \(\nu\) are isomorphisms.
\end{lem}
\begin{proof}
  Let \(\tau = \mu \nu\) be an automorphism. Now, replace \(\mu\) with
  \(\tau^{-1} \mu\). Then, we have split short exact sequence \[
    \begin{tikzcd}
      0 \rar & M \ar[r,swap,"\nu"] & N \rar \ar[l,bend right = 30, swap,"\mu"] & N/M \rar & 0
    \end{tikzcd}
  \]
  and so \(N \isom M \oplus N/M\). However, \(N\) is indecomposable,
  so \(N = M\).
\end{proof}
\begin{lem}
  Let \(0 \neq M\) be indecomposable, Artinian, and Noetherian. Let \(\tau_1, \ldots, \tau_r \in \End_R M\) satisfy \(\tau_1 +
  \cdots + \tau_r \in \Aut_R (M)\). Then, at least one \(\tau_i\) is
  an automorphism of \(M\)
\end{lem}
\begin{proof}
  Since \(\tau = \tau_1 + (\tau_2 + \cdots + \tau_r)\), it suffices to
  prove for the sum of 2 endomorphisms. Let \(\psi = \tau_1 + \tau_2\)
  be an automorphism. Then, \(\phi_1 = \psi^{-1} \tau_1\) and \(\phi_2
  = \psi^{-1} \tau_2\). We then have \(\phi_1 + \phi_2 = 1_M\) and so,
  since \(\phi_1 = 1_M - \phi_2\), we get \(\phi_1 \phi_2 = \phi_2
  \phi_1\). Now, we can apply the binomial theorem to get \[
    0 \neq 1_M^m = (\phi_1 + \phi_2)^m = \sum_{k=1}^m \binom{m}{k} \phi_1^k \phi_2^{m-k}
  \]
  So, at least one of the \(\phi_i\)'s cannot be nilpotent. Without
  loss of generality, assume \(\phi_1\) is not nilpotent. Since
  \(\phi_1\) is not nilpotent and \(M\) has a composition series, we
  can apply Fitting's Lemma to get \[
    M \isom \phi_1^\infty(M) \oplus \phi_1^{-\infty}(M)
  \]
  However, \(M\) is indecomposable and \(\phi_1^\infty(M) \neq 0\), so
  \(M = \phi_1^\infty(M)\) and \(\ker \phi_1 \subset
  \phi_1^{-\infty}(M) = 0\), so \(\phi_1\) is an
  automorphism. Therefore, \(\tau_1\) is an automorphism.
\end{proof}
We are now ready to prove the Krull-Schmidt theorem.
\begin{proof}[Proof of Krull-Schmidt]
  Let \(0 \neq M\) be an Artinian and Noetherian \(R\)-module. Suppose
  \begin{align*}
    M & = M_1 \oplus \cdots \oplus M_h \\
    & = N_1 \oplus \cdots \oplus N_k
  \end{align*}
  where the \(M_i,N_j\) are indecomposable. We will proceed by
  induction on \(h\). If \(h=1\), then \(M\) is already indecomposable
  and we are done. Now, for the inductive step, let \(\mu \from M \to
  M_i\) be the projection onto the \(i\)th summand and, similarly, let
  \(\nu_j \from M \to N_j\) be a projection. Then, we have \(\mu_i
  \mu_j = \delta_{ij} \mu_i\) and \(\nu_i \nu_j = \delta_{ij}
  \nu_j\). Now, we have the following identity \[
    \mu_1 = \mu_1 \circ 1_M = \mu_1(\nu_1 + \cdots + \nu_k)
  \]
  Now, let \(\ov{\nu}_j = \nu_j |_{M_i} \from M_i \to M\). If we
  restrict the above identity to \(M_1\), we get \[
    1_{M_1} = \mu_1 \ov{\nu}_1 + \cdots + \mu_1 \ov{\nu}_k \in \Aut M_1
  \]
  Thus, we can use the lemma above to assume that \(\mu_1 \ov{nu}_1
  \in \Aut M_1\). Furthermore, by the other lemma, since \(\mu_1 \ov{\nu}_1\) is
  an automorphism, we get that \(\ov{nu}_1\) is an isomorphism, as is
  \(\mu_1 \from N_1 \to M_1\). Now,
  consider \[
    N_1 + M_2 + \cdots + M_n \submodule M
  \]
  We wish to show this sum is actually direct. Assume there is \(n_1
  \in N_1\) and \(m_2 \in M_2, \ldots, m_h \in M_h\) such that \[
    n_1 + m_2 + \cdots + m_h = 0.
  \]
  We can just apply \(\mu_1\) to this sum above to get \[
    0 = \mu_1(n_1 + m_2 + \cdots + m_h) = \mu_1(n_1)
  \]
  However, \(\mu_1\) is an isomorphism between \(N_1\) and \(M_1\), so
  \(\mu_1(n_1) = 0 \implies n_1 = 0\). So, furthermore, it must be
  that \(m_2 = \cdots = m_h = 0\) by the direct sum construction,
  so \[
    M' := N_1 + M_2 + \cdots + M_h = N_1 \oplus M_2 \oplus \cdots
    \oplus M_h \submodule M
  \]
  Thus, we wish to finally show that \(M' = M\). Let \[
    \rho \from M_1 \oplus \cdots \oplus M_h \to M'
  \]
  be given by \(\rho = \ov{\nu}_1 + \mu_2 + \cdots + \mu_h\). Note
  that \(\rho\) is an isomorphism since \(\ov{\nu}_1\) is an
  isomorphism from \(M_1 \to M\). Since
  \(M\) is Artinian, there is some \(a \gg 0\) such that \[
    \rho^{a+1}(M) = \rho^a(M)
  \]
  Hence, given \(m \in M\), there is some \(m' \in M\) such that
  \begin{align*}
    \rho^{a+1}(m') = \rho^a(m)
    & \implies \rho^a(m-\rho(m')) = 0 \\
    & \implies m-\rho(m') \in \ker \rho^a = \{0\}
  \end{align*}
  Thus, \(M' = M\) and \(\rho\) is an automorphism of
  \(M\). Furthermore, \(\rho(M_i) = M_i\) for \(i = 2,\ldots,h\),
  \(\rho(M_1) = N_1\), and \[
    N_1 \oplus M_2 \oplus \cdots \oplus M_h = \rho(M) = M = M_1 \oplus
    \cdots \oplus M_h
  \]
  Hence,
  \begin{align*}
    M_2 \oplus \cdots \oplus M_h
    & \isom M/M_1 \\
    & \isom \rho(M)/\rho(M_1) \\
    & \isom \rho(M)/N_1 \\
    & \isom (N_1 \oplus \cdots \oplus N_k) / N_1 \\
    & \isom N_2 \oplus \cdots \oplus N_k
  \end{align*}
  Thus, by induction, \(h-1 = k-1 \implies h = k\) and \(M_i \isom
  N_i\) for \(i=2,\ldots,h\), and we already know \(M_1 \isom N_1\) by
  above.
\end{proof}
\begin{cor}
  Let \(N\) be a direct summand of \(M\) as in the Krull-Schmidt
  theorem. Then, there exist a subset of \(\{M_i\}\) indecomposable submodules 
  \(M_{i_1}, \cdots , M_{i_t}\) such that \[
    N \isom M_{i_1} \oplus \cdots \oplus M_{i_t}
  \]
\end{cor}
\section{Nakayama's Lemma: A Hard to Remember Lemma}
While one statement of Nakayama's Lemma may be easy to remember, there
are actually many special cases of this lemma and so it has a
reputation of being hard to remember. To understand and prove
Nakayama's Lemma, we must first lay the groundwork of Jacobson
radicals.
\begin{defn}
  The \de{Jacobson radical} (often just called the radical) \(R\) of a
  ring \(A\) is \[
    R = \{x \in A \st xM = 0, \forall \text{ simple }A\text{-modules }M\}
  \]
\end{defn}
\begin{defn}
  A ring \(A\) is called \de{semisimple} when its radical \(R = 0\). 
\end{defn}
\begin{thm}\label{radical-is-2-sided-ideal}
  The radical \(R\) of \(A\) is a two-sided ideal of \(A\) and \(A/R\)
  is semisimple.
\end{thm}
\begin{proof}
  If \(M\) is a simple \(A\)-module, then \(M\) is also an
  \(A/R\)-module. Conversely, any \(A/R\)-modules are naturally
  \(A\)-modules. Thus, if there were an \(x + R \in A/R\) such that
  \((x+R)M = xM = 
  0\) for all simple \(A/R\)-modules \(M\), then \(x\) would also
  annihilate all simple \(A\)-modules and thus \(x \in R\), so \(x+R =
  0 + R\) and thus the radical of \(A/R\) is trivial.
\end{proof}
\begin{thm}\label{radical-is-intersection-of-maxl-ideals}
  Let \(R\) be the radical of a ring \(A\). Then, \[
    R = \Intersect_{M \text{ simple }} \Ann_A (M) = \Intersect_{\substack{I
      \ideal A \\ \text{Maximal left ideal}}} I
  \]
  That is, \(R\) is the intersection of all maximal left ideals of \(A\).
\end{thm}
\begin{proof}
  Let \(x \in R\) and let \(I\) be a maximal left ideal. We want to
  show that \(x \in I\). Since \(I\) is maximal, \(A/I\) is a simple
  left \(A\)-module. thus, \(x \in \Ann(A/I)\). Thus,
  \begin{align*}
    0 = x(1+I) & \implies x \in I \\
               & \implies x \in \Intersect I \\
               & \implies R \subset \Intersect I
  \end{align*}
  Conversely, let \(x \in \Intersect I\). Let \(M\) be a simple module
  \(0 \neq m \in M\). We show \(xm = 0\). We have
  \begin{align*}
    A & \overset{\theta}{\onto} M \\
    a & \mapsto am
  \end{align*}
  Then, \(M \isom A/I\) where \(I = \Ann_A(m)\). Thus, \(xm = 0\).
\end{proof}
\begin{thm}\label{one-minus-radical-is-unit}
  Given a ring \(A\), \[
    \rad A = \{x \in A \st 1-axb \in A^\times, \forall a,b \in A\}
  \]
  where \(A^\times\) is the group of units of \(A\).
\end{thm}
\begin{proof}
  Let \(x \in \rad A\) with \(a,b \in A\). Then, \(\rad A \ideal A\)
  by \ref{radical-is-2-sided-ideal}, so \(axb \in \rad A\). Now, write
  \(y := axb\). We wish to show \(1-y\) is a unit. \\

  If \(A(1-y) \neq A\), let \(I\) be a maximal left ideal containing
  \(A(1-y)\). Then, by \ref{radical-is-intersection-of-maxl-ideals},
  \(\rad A \subset I\), so \(y \in \rad A \subset I\) and \(1-y \in
  A\). This would give that \(1 = y + (1-y) \in I \implies I = A\),
  which is a contradiction of the maximality of \(I\). So, \(A(1-y) =
  A\) for all \(y \in \rad A\). \\

  Now, since \(A(1-y) = A\), there is a \(t\) such that \(t(1-y) =
  1\). So, we wish to show \((1-y)t = 1\). Rearranging our expression
  \(t(1-y) = 1\), we get
  \begin{align*}
    1-t = -ty \in \rad A
    & \implies A(1-(1-t)) = A
    & \text{ by above since }A(1-y) = A, \forall y \in \rad A \\
    & \implies At = A \\
    & \implies \exists u \in A \text{ such that } ut=1 \\
    & \implies u = ut(1-y) = 1-y
    & \text{since }t(1-y)=1 \text{ and }ut=1 \\
    & \implies 1-y \in A^\times
  \end{align*}
  Thus, we have shown one containment. \\

  Now, let \(x \in A\) be such that \(1-axb \in A^\times\) for all
  \(a,b \in A\). We want to prove \(x \in \rad A\). We show \(xM = 0\)
  if \(M\) is a simple \(A\)-module. Let \(0 \neq m \in M\). Assume
  \(xm \neq 0\). Then, \(M = Axm\) since \(M\) is simple so \((xm) =
  M\). Thus, \(m = axm\) for some \(a \in A\) and so \((1-ax)m =
  0\). However, \(1-ax \in A^\times\), so \(m = 0\), which is a
  contradiction. Thus, \(xM = 0\) and so \(x \in \rad A\).
\end{proof}
\begin{thm}[Nakayama's Lemma]\label{nakayamas-lemma}
  \cite{jacobson}*{p 415} Let \(A\) be a ring and let \(M\) be a finitely-generated
  \(A\)-module. If \((\rad A)M = M\), then \(M = 0\).
\end{thm}
\begin{proof}
  Let \(m_1, m_2, \ldots, m_k\) be a minimum generating set of
  \(M\). Given \(m \in M\), there exists \(a_1, \ldots, a_k \in A\)
  such that \[
    m = a_1 m_1 + \cdots + a_k m_k
  \]
  Take \(m = m_1\). Then, \[
    (1-a_1)m_1 = a_2 m_2 + \cdots + a_k m_k
  \]
  If \((\rad A) M = M\), we can assume each \(a_i \in \rad A\), we can
  assume each \(a_i \in \rad A\) since \(\rad A\) is a 2-sided ideal
  by theorem \ref{radical-is-2-sided-ideal}. Thus, by
  \ref{one-minus-radical-is-unit}, \(1-a_1\) is a unit, so \[
    m_1 = (1-a_1)^{-1} a_2 m_2 + \cdots + (1-a_2)^{-1} a_k m_k \in
    \langle m_2, \ldots, m_k \rangle
  \]
  However, this contradicts the minimality of the the generating set.
\end{proof}
\begin{cor}
  Let \(M\) be a finitely generated left \(A\)-module and let \(N
  \submodule M\) be a submodule such that \[
    N+(\rad A)M = M
  \]
  Then, \(N = M\)
\end{cor}
\begin{proof}
  Given the above, \(\rad A(M/N) = M/N\) simply by quotienting both
  sides of the equality by \(N\). Thus, by Nakayama's lemma, \(M/N =
  0\) and so \(M=N\).
\end{proof}
\begin{cor}
  Let \(J\) be a maximal ideal in a ring \(A\). Then, \(\rad A \subset
  J\).
\end{cor}
\begin{proof}
  Assume not. Then \(\rad A + J = A \implies J = A\) by the corollary
  above, which is a contradiction to \(J\) being a maximal ideal.
\end{proof}
\section{Completely Reducible Modules}
Understanding all the modules of a ring, in general, is an incredibly
difficult problem. However, usually the first step in such a program
is understanding the simple (or irreducible) modules. Understanding
these modules would then give a complete understanding of completely
reducible modules.
\begin{defn}
  Let \(A\) be a ring. A left \(A\)-module \(M\) is called
  \de{completely reducible} if, given any submodule \(N \subset M\),
  there exists a submodule \(N'\) such that \[
    M = N \oplus N'
  \]
\end{defn}
However, there are equivalent notions of completely reducible.
\begin{thm}\label{cr-module-equivs}
  Assume \(M\) is a left \(A\)-module. Then, the following are
  equivalent.
  \begin{enumerate}
  \item \(M\) is completely reducible.
  \item \(M\) is a direct sum of irreducible submodules, that is \[
      M \isom \bigoplus_{i \in I} L_i
    \]
    where \(I\) is some indexing set and \(L_i\) is simple.
  \item \(M\) is a sum of irreducible submodules, that is \[
      M \isom \sum_{i \in I} L_i
    \]
    where \(I\) is some indexing set and \(L_i\) is simple.
  \end{enumerate}
\end{thm}
To prove this result, we use the following lemma.
\begin{lem}
  If an \(A\)-module \(M\) is completely reducible, so is any
  submodule \(N \submodule M\).
\end{lem}
\begin{proof}[Proof of Lemma]
  Consider submodule \(N\) of completely reducible a \(A\)-module
  \(M\). Then, by definition, there exists an \(N'\) such that \[
    M \isom N \oplus N'.
  \]
  Now, let \(S \submodule N\). Then, we also have \(S \submodule M\)
  and so there is a submodule \(T\) such that \[
    M \isom S \oplus T.
  \]
  Thus, we have
  \begin{align*}
    N & = (S \oplus T) \intersect N \\
      & = S \oplus (T \intersect N)
  \end{align*}
  since \(S \submodule N\). Since \(T \intersect N \submodule N\),
  then \(N\) is completely irreducible.
\end{proof}
\begin{lem}
  Let \(M\) be a completely reducible \(A\)-module. Then, every
  nontrivial submodule \(N\) of \(M\) contains an irreducible submodule.
\end{lem}
\begin{proof}[Proof of Lemma]
    Let \(N \submodule M\) and \(0 \neq n \in N\). Consider the
    collection \[
      S = \{N' \submodule N \st n \not \in N'\}
    \]
    We note that \(S\) is nonempty since \((0) \in S\). Thus, since
    \(S\) is nonempty, it must contain a maximal element, say
    \(N_0\). By the lemma above, we know that \(N\) is completely
    reducible and so \(N = N_0 \oplus N_1\) for some submodule \(N_1
    \submodule N\). \(N_1\) must be irreducible because, if not, then
    there would be a proper submodule \(N_2 \submodule N_1\) and \(N_1
    = N_2 \oplus N_3\) for some submodule \(N_3 \submodule
    N_1\). However, this would give us \[
      N = N_0 \oplus N_2 \oplus N_3 \implies \text{ either } n \not
      \in N_0 + N_2 \text{ or } n \not \in N_0 + N_3
    \]
    since \((N_0 + N_2) \intersect (N_0 + N_3) = N_0\). Such a result
    contradicts the maximality of \(N_0\) in \(S\), and so it must be
    that \(N_1\) is irreducible.
\end{proof}
\begin{proof}[Proof of Theorem]
  We first assume that \(M \neq (0)\), otherwise the theorem is
  immediately true.
  \begin{enumerate}
  \item[(\((a) \implies (b)\)).] Let \(\{M_i \st i \in I\}\) be the
    collection of all irreducible submodules of \(M\) and let \[
      T = \left\{J \subset I \ \st \ \sum_{j \in J} M_j \text{
          is direct} \right\}
    \]
    \(T\) is nonempty since it contains at least singleton sets are in
    \(T\) and any union
    of an ascending chain of elements in \(T\) is in \(T\). Thus, we
    can apply Zorn's lemma to get a maximal element of \(T\), say
    \(J_0\). Now, let \[
      M' := \bigoplus_{j \in J_0} M_j
    \]
    be properly contained in \(M\). Then, by complete reducibility of
    \(M\), we get \[
      M = M' \oplus M''
    \]
    for \((0) \neq M'' \submodule M\). However, 
    since \(M''\) must be completely reducible, by the lemma above,
    \(M''\) must contain an an irreducible submodule, say
    \(M_{i_0}\). Then, \[
      \left(\bigoplus_{j \in J_0} M_j \right) + M_{i_0} \text{ is direct}
    \]
    and thus we have violated the maximality of \(J_0\).
  \item[(\((b) \implies (c)\)).] This result is immediate by the
    definition of direct sum.
  \item[(\((c) \implies (a)\)).] Let \(N \submodule M\) and let \(N'\)
    be a maximal submodule with respect to the property that \(N
    \intersect N' = (0)\). Assume \(M \neq N \oplus N'\). Then, there
    is an \(m \in M, m \not \in N \oplus N'\). However, since \(M\) is
    a sum of irreducible submodules, \(m = m_1 + \cdots + m_k\) where
    each \(m_i\) belonds to an irreducible summand. Thus, at least one
    \(m_i \not \in N \oplus N'\) since \(m \not \in N \oplus N'\) and
    thus \(M_i \not \subset N \oplus N'\). However, \(M_i\) is
    irreducible and so \(M_i \intersect (N \oplus N') = (0)\). Thus,
    \(N' \propsubgroup N' + M_i\) and \((N'+M_i) \intersect N = (0)\),
    thus violating the maximality of \(N'\). Thus, \(M = N \oplus N'\).
  \end{enumerate}
\end{proof}
\begin{prop}
  Let \(M\) be a completely reducible module for a ring \(A\). Then,
  \(M\) satisfies the ascending chain condition if and only if \(M\)
  satisfies the descending chain condition.
\end{prop}
\section{Nilpotent and non-Nilpotent Ideals in Artinian Rings}
To get to the Artin-Wedderburn Theorem, we must have an understanding
of idempotents in Artinian rings, which are intimately (non)-related to
nilpotent ideals. We wish to culminate in a theorem that says any
non-nilpotent left ideal in a left-artinian ring must contain an
idempotent. First, we present a theorem similar to Schur's Lemma, but
for Noetherian rings.
\begin{thm}
  Let \(M\) be a (left) Noetherian \(A\)-module for ring \(A\). If \(f
  \in \End_A(M)\) is surjective, then \(f\) is an isomorphism
\end{thm}
\begin{proof}
  If \(f\) is surjective, then \(\im f = M\) and so \(f\) is not
  nilpotent. Thus, by Fitting's Lemma \[
    M = f^\infty(M) \oplus f^{-\infty}(M)
  \]
  we seek to show \(\ker f = 0\). For some integer \(n\), \(\ker f^n =
  \ker f^{n+1}\). Let \(x \in \ker f\). Then, since \(f\) is
  surjective, \[
    f^n(M) = f^{n+1}(M) = M
  \]
  Thus, \(x = f^n(y)\) for some \(y \in M\). We then see \[
    0 = f(x) = f^{n+1}(y) \implies y \in \ker f^{n+1} = \ker f^n
    \implies 0 = f^n(y) = x
  \]
  So, we have that \(\ker f = 0\) and thus \(f\) is injective.
\end{proof}
\begin{thm}
  Let \(A\) be a left Noetherian ring. If \(a,b \in A\) be such that \(ab =
  1\), then \(ba = 1\) and \(a,b \in A^\times\).
\end{thm}
\begin{proof}
  Let \(M = {}_A A\). Then, \(M\) is a (left) Noetherian
  \(A\)-module. Then, since \(ab = 1\), \[
    A = Aab \subset Ab \subset A
  \]
  and so \(Ab = A\). Thus, we have a map
  \begin{align*}
    f \from M & \onto M \\
    x & \mapsto xb
  \end{align*}
  and so, by the theorem above, \(f\) is an isomorphism. However,
  \begin{align*}
    f(1-ba) & = (1-ba)b \\
            & = b-bab\\
            & = b-b & \text{since }ab = 1\\
    & = 0
  \end{align*}
  Thus, \(1-ba = 0 \implies ba = 1 = ab\). 
\end{proof}
\begin{defn}
  An \de{idempotent} of a ring \(A\) is an element \(0 \neq e \in A\)
  such that \(e^2 = e\).
\end{defn}
One advantage of idempotents is that they allow us to ``project'' the
ring onto ``orthogonal'' compotents, that is, given a ring \(A\) with
idempotent \(e\), then, as a module over itself \[
  {}_A A \isom {}_A Ae \oplus {}_A A(1-e)
\]
This also tells us the following
\begin{rmk}
  \({}_A Ae\) is projective as an \(A\)-module since it is the direct
  summand of a free \(A\)-module.
\end{rmk}

\begin{defn}
  We call a left or right ideal, \(I\), \de{nilpotent} if there is an \(m
  \in \N\) such that \(I^m = \{0\}\).
\end{defn}
\begin{example}
  Consider \(R = \Z/p^n \Z\) where \(p\) is a prime number. Then,
  since \(R\) is a PID, every proper ideal is generated by some \(p^k
  + (p^n)\) and is nilpotent. The only idempotent of \(R\) is \(1 + (p^n)\).
\end{example}
\begin{thm}
  Let \(N\) be a nilpotent left ideal of a ring \(A\). Let \(x \in A\)
  be non-nilpotent such that \(x^2-x \in N\). Then, the left ideal
  \(Ax\) has an idempotent \(y\).
\end{thm}
The idea to proving this theorem is to take \(A\) and factor out
\(N\). Then, one can find an idempotent \(y\) such that, under the quotient
map \(q \from A \to A/N\), \(q(x) = q(y)\).
\begin{proof}
  Assume \(N^k = 0\) for some positive integer \(k\). Let \(m_1 :=
  x^2-x \in N\). If \(m_1 = 0\), then \(x^2-x = 0\) so \(x\) is an
  idempotent itself and we can take \(y = x\). \\

  Assume \(m_1 \neq 0\). Then, let \[
    x_1 := x + m_1 - 2xm_1 \in Ax \text{ since }m_1 \in Ax
  \]
  Note that \(x_1,x,m_1\) all commute. Then, note that \(x_1\) is not nilpotent
  as well and \(x^2 = x + m_1\). Consider
  \begin{align*}
    x_1^2 - x_1
    & = (x+m_1-2xm_1)^2 - (x+m_1-2xm_1) \\
    & = x^2 + xm_1 - 2x^2m_1 + m_1 x + m_1^2 -2xm_1^2 - 2x^2m_1 -
      2xm_1^2 + 4 x^2 m_1^2 - x - m_1 + 2x m_1 \\
    & = 4x^2 m_1^2 - 4x^2 m_1 - 4x m_1^2 + x^2 + 4xm_1 + m_1^2 - x - m_1 \\
    & = 4x^2 m_1^2 - 4x^2 m_1 - 4x m_1^2 + 4xm_1 + m_1^2 \\
    & = (4x^2-4x)m_1^2 + (-4x^2+4x)m_1 + m_1^2 \\
    & = (4x^2-4x-4m_1)m_1^2 + (-4x^2+4x+4m_1)m_1 + m_1^2 + 4m_1^3 - 4m_1^2 \\
    & = 4m_1^3 - 3m_1^2
  \end{align*}
  Then, take \[
    m_2 = 4m_1^3 - 3m_1^2 \ x_2 = x_1 + m_2 - 2x_1 m_2.
  \]
  and note that \(m_2\) contains \(m_1^2\) as a factor. Thus, we can
  successively construct non-nilpotent elements \(x_1, x_2, \ldots\)
  in \(Ax\) such that \(x_i^2 - x_i\) contains \(m_1^{2^i}\) as a
  factor and commutes with \(x\). Since \(m_1\) is nilpotent, then
  \(m_1^{2^i} = 0\) for sufficiently large \(i\) and so \(x_i^2 - x_i
  = 0\) for some sufficiently large \(i\). Therefore, for that
  sufficiently large \(i\), \(x_i\) is nilpotent.
  \end{proof}
\begin{rmk}
  Note that any nilpotent ideal cannot contain an idempotent
  element. However, the following theorem gives us a (useful!)
  converse to that fact.
\end{rmk}
\begin{thm}\label{artinian-non-nilp-ideal-has-idem}
  Let \(L\) be a non-nilpotent left ideal in a left-artinian ring
  \(A\). Then, \(L\) contains an idempotent \(e\).
\end{thm}
\begin{proof}
  We seek to use the theorem above to get such an idempotent by
  finding a non-nilpotent \(x \in A\) such that \(x^2-x\) is in a
  nilpotent ideal. \\

  Choose a minimal left ideal \(L_1 \subset L\) which is not
  nilpotent. Then, \(0 \neq L_1 L_1 \subset L_1\) is not nilpotent, so
  \(L_1 L_1 = L_1\) by the minimality of \(L_1\). \\

  Let \(I\) be a left ideal contained in \(L_1\) such that \(L_1 I
  \neq 0\) and minimal with respect to this property. Let \(a \in I\)
  be such that \(L_1 a \neq 0\). Then, \[
    L_1 L_1 a = L_1^2 a = L_1 a \neq 0 \text{ and } L_1 a \subset I
  \]
  Hence, \(I = L_1 a\) by the minimality of \(I\). Thus, there is an
  \(x \in L_1\) such that \(a=xa\). Hence, \[
    0 \neq a = xa = x^2 a = \cdots = x^k a = \cdots
  \]
  Therefore, \(x\) is not nilpotent. \\

  Let \(N = \{b \in L_1 \st ba = 0\}\). This is a left ideal contained
  in
  \(L_1\). Since \(xa = x^2 a\), \[
    (x-x^2)a = 0 \implies x-x^2 \in N
  \]
  Also, \(L_1 a \neq 0\) so \(N \propsubset L_1\). Hence, \(N\) is
  nilpotent by the minimality of \(L_1\) as a non-nilpotent
  ideal. Thus, we now have \(x,N\) as in the theorem above and so
  there is an idempotent \(e \in Ax \subset L_1 \subset L\).
\end{proof}
\begin{thm}\label{for-artinian-rad-is-largest-nilp-ideal}
  Let \(A\) be a left artinian ring. Then \(\rad A\) is the largest
  nilpotent left ideal.
\end{thm}
\begin{proof}
  
\end{proof}
\begin{prop}
  If \(f \from A \to B\) is a surjective homomorphism of rings,
  then \[
    f(\rad A) \subset \rad B
  \]
\end{prop}
\section{The Radical of an Artinian Ring}
In this section, we seek to understand the radical or an artinian ring \(A\)
and show that when it is trivial, we gain a huge amount of
insight into \(A\)-modules. Namely, for semisimple \(A\), every \(A\)-module is
completely reducible and every irreducible \(A\)-module is
isomorphic to a non-zero minimum left ideal of \(A\). Perhaps most
surprisingly of all, this tells us that a semisimple \(A\) has only a
finite number of irreducible modules, which is not a typical
phenemenon! We start by finding a deeper insight into the relationship
between \(\rad A\) and the nilpotent ideals of \(A\).
\begin{thm}
  Let \(A\) be a left artinian ring. Then, the sum of all left nilpotent
  ideals of \(A\) is a nilpotent ideal, say \(N\). It contains every
  nilpotent right ideal of \(A\). Also, the quotient ring \(A/N\) has
  no nontrivial nilpotent ideal.
\end{thm}
To show this, we use the following lemma
\begin{lem}
  Any finite sum of nilpotent ideals is nilpotent.
\end{lem}
\begin{proof}[Proof of Lemma]
  If we assume the result is true for a sum of \(r-1\) nilpotent
  ideals, then it is true for \(r\) nilpotent ideals. Consider
  nilpotent ideals \(N_1, \ldots, N_r\). Then, \[
    N_1 + \cdots + N_r = \underbrace{(N_1 + \cdots +
      N_{r-1})}_{\text{Nilpotent ideal by assumption}} + N_r \text{ is
    nilpotent.}
  \]
  Thus, it suffices to show the result is true for the sum of two
  nilpotent ideals. \\

  Let \(N_1, N_2\) be nilpotent left ideals. Then, there exist
  positive integers \(q,r\) such that \[
    N_1^q = N_2^r = 0.
  \]
  And element in \((N_1 + N_2)^{q+r}\) is a finite sum of products \[
    y_1 y_2 \cdots y_{q+r}, \ y_i \in N_1 \union N_2, s \in \N
  \]
  We can assume at least \(q\) of the \(y_i\) belong to \(N_1\) or at
  least \(r\) of the \(y_i\) belong to \(N_2\). Without loss of
  generality, we assume the
  former. Then, the product above may be written \[
    (x_1 x_2 \cdots x_{i_1})(x_{i_1 + 1} \cdots x_{i_2}) \cdots
    (x_{i_{s-1}+1} \cdots x_{i_s}) \cdots, 
  \]
  where \(x_{i_1}, x_{i_2}, \ldots,
  x_{i_s} \in N_1\) and \(s \geq q\). Since \(N_1\) is a left ideal,
  each ``segment'' belongs to \(N_1\). Thus, \[
    y_1 y_2 \cdots y_{q+r} \in N_1^q = 0  \implies y_1 y_2 \cdots
    y_{q+r} = 0
  \]
  However, \(y_1 y_2 \cdots y_{q+r}\) was an arbitrary element of
  \((N_1 + N_2)^{q+r}\), so \((N_1 + N_2)^{q+r} = 0\) and thus the sum
  of 2 nilpotent (left) ideals is nilpotent.
\end{proof}
\begin{proof}[Proof of Theorem]
  Let \(\{N_i\}\) be a set of nilpotent left ideals. Then, the
  elements of \(N\) are finite sums, i.e. \[
    n \in N \implies n = x_1 + \cdots + x_r, x_i \in \text{ some }N_i
  \]
  Since the sum of left ideals is a left ideal, \(N\) is a left
  ideal. \\

  Assume \(N\) is not nilpotent, then since \(A\) is left-artinian,
  \(N\) contains an idempotent, \(e\)
  (see \ref{artinian-non-nilp-ideal-has-idem}). Thus, there is a finite sum
  of nilpotent ideals such that \[
    e \in N_1 + \cdots + N_r
  \]
  However, \(N_1 + \cdots + N_r\) is nilpotent by the lemma above, and
  so there is a \(k\) such that \[
    e = e^k \in (N_1 + \cdots + N_r)^k = 0
  \]
  which is a contradiction. \\

  If \(N_i\) is a nilpotent left ideal, then \(N_i A\) is a nilpotent
  (two-sided) ideal since, for large enough \(t \in \N\), \[
    (N_i A)^t = N_i(A N_i) \cdots (A N_i) A \subset N_i (N_i)^{t-1}A = 0
  \]
  Thus, \(N\) is an ideal. \todo{How does this follow exactly?} If
  \(I\) is a nilpotent right ideal, \(AI\) is a nilpotent left ideal
  by the same argument. Thus, \(I \subset AI \subset N\). \\

  Since \(N\) is a two-sided ideal, then every left ideal of \(R/N\)
  is of the form \(I/N\) for some left ideal \(I\) of \(R\), by the
  correspondance theorem. Thus, \[
    (I/N)^k = 0 \iff I^k \subset N \iff I
    \text{ is nilpotent} \iff I \subset N
  \]
  since \(N\) is the sum of all nilpotent left ideals. Thus, \(R/N\)
  contains no nilpotent left ideal except \(0\), and by the discussion
  above, also contains no nontrivial nilpotent right or two-sided ideals.
\end{proof}
\begin{thm}
  Let \(A\) be a left artinian ring. If is \(N\) the sum of all
  left nilpotent ideals, then \(N = \rad A\).
  \begin{proof}
    We already have that \(N \subset \rad A\) by
    (\ref{for-artinian-rad-is-largest-nilp-ideal} \todo{Is this not
      the same result as the theorem we are proving? Did we show
      something weaker, earlier?}) since \(N\) is
    nilpotent by the above theorem. However, consider the sequence of
    ideals \[
      \rad A \supset (\rad A)^2 \supset \cdots \supset (\rad A)^n
      \supset \cdots
    \]
    Since \(A\) is artinian, there is some \(k\) such that \[
      (\rad A)^k = (\rad A)^{k+1} = \cdots
    \]
    and thus \((\rad A)(\rad A)^n = (\rad A)^n\). Thus, by Nakayama's
    lemma (\ref{nakayamas-lemma}), \((\rad A)^n = 0\). Thus, \(\rad
    A\) is nilpotent.
  \end{proof}
\end{thm}
\begin{thm}
  Let \(A\) be a left artinian ring. Then, \({}_A A\) is completely
  reducible if and only if \(A\) is semisimple.
\end{thm}
\begin{proof}
  \((\impliedby).\) Let \(A\) be semisimple. Then, \(\rad A =
  0\). Now, Let \(I_1\) be minimal among nontrivial non-nilpotent left
  ideals of \(A\). Then, \(I_1\) has an idempotent, say \(e_1\). Thus,
  \(e_1 \in I_1 \implies Ae_1 \subset I_1\), but \(I_1\) is minimal
  among non-nilpotent ideals, so \(I_1 = Ae_1\). Now, notice that \[
    A = Ae_1 \oplus A(1-e_1) \implies {}_A A = I_1 \oplus I_1', \
  \]
   where \(I_1' = A(1-e_1)\) is a non-nilpotent left ideal if \(I_1'
   \neq 0\). Then, we can repeat this process with \(I_1'\) \todo{Why
     is \(I_1'\) minimal?} to get \[
     {}_A A = I_1 \oplus I_2 \oplus I_2', \ I_2 = Ae_2
   \]
   \todo{There is more work to do here.} Continuing, we get \[
     {}_A A = I_1 \oplus I_2 \oplus \cdots \oplus I_r
   \]
   where \(I_i = Ae_i, e_i\) idempotent. Each \(I_i\) is irreducible
   as a left \(A\)-module since \(A\) is semisimple. \\

   \((\implies)\). Let \({}_A A\) be completely reducible. Consider
   that \(N = \rad A\) is a left ideal of \(A\) and thus also a
   submodule of \({}_A A\) and thus \[
     {}_A A = N \oplus N'
   \]
   for some left ideal \(N'\) of \(A\). Then, \[
     1 = x + x'
   \]
   for some \(x \in N, x' \in N'\), which then yields \[
     x \cdot 1 = x(x+x') = x^2 + xx' \implies x - x^2 = x x' \in N
     \intersect N' = 0
   \]
   Thus, \(x-x^2 = 0 \implies x = x^2 = \cdots = 0\) since \(N\) is
   nilpotent. Thus, \(x' = 1\) and \(N' = A \implies \rad A = N = 0\)
   and thus \(A\) is semisimple.
\end{proof}
\begin{thm}
  Let \(A\) be a (left) artinian ring. Then, \(A\) is semisimple if
  and only if every \(A\)-module is completely reducible.
\end{thm}
\begin{proof}
  To prove \((\impliedby)\), we simply note that \(_{A} A\) is an
  \(A\)-module, and so the previous theorem tells us that \(A\) is
  semisimple. \\

  For \((\implies)\), assume \(A\) is semisimple and \(M\) is an
  \(A\)-module. Then, it suffices to show \(M\) is a (not-necessarily
  direct) sum of irreducible \(A\)-modules by
  \ref{cr-module-equivs}. By the theorem above, we know \[
    {}_A A \isom L_1 \oplus \cdots \oplus L_n
  \]
  Now, given \(0 \neq m \in M\), consider \(\phi \from A \to M\) given
  by \(a \mapsto a.m\). Then, \(\phi|_{L_i} \from L_i \to M\) is
  either injective or the \(0\) map. Thus, we get \[
    A.M = M \implies (L_1 \oplus \cdots \oplus L_n).M = M \implies
    \phi(L_1) + \cdots + \phi(L_n) = M
  \]
  but since each \(\phi(L_i)\) is irreducible or \(0\), it must be
  that \(M\) is completely reducible.
\end{proof}
\begin{thm}
  Let \(A\) be artinian and semisimple. Then, every irreducible
  \(A\)-module \(L\) is isomorphic to a non-zero minimum left ideal of
  \(A\). Thus, \(A\) has only a finite number of irreducible modules.
\end{thm}
\begin{proof}
  Similar to above, we know \[
    {}_A A \isom L_1 \oplus \cdots \oplus L_n
  \]
  and, for irreducible \(A\)-module \(M\) and \(0 \neq m \in M\), we have a map \(\phi \from
  A \to M\) given by \(a \mapsto a.m\). Then, \[
    M = \phi(L_1) + \cdots + \phi(L_n)
  \]
  and one of the summands must be nonzero, say \(\phi(L_i)\). However,
  since \(M\) is irreducible, this means that \(M = \phi(L_i)\) and
  thus \(\phi|_{L_i}\) is an isomorphism. Since \(A\) has only a
  finite number of ideals up to isomorphism by the Krull-Schmidt
  theorem, this means that \(A\) has 
  only a finite number of irreducible modules. 
  % Then,
  % \(L_i\) contains an idempotent \(e_i\) by
  % \ref{artinian-non-nilp-ideal-has-idem} and since \(L_i\) is
  % irreducible, \(L_i = Ae_i\). 
\end{proof}
\begin{thm}
  Let \(A\) be artinian and semisimple. Let \(Ae\) be a left ideal for
  idempotent \(e\). Then, \(Ae\) is irreducible if and only if \(eAe\)
  is a division ring.
\end{thm}
\begin{proof}
  Let us assume \(Ae\) is irreducible and let \(L\) be any nonzero
  left ideal of \(eAe\). Then, \[
    AL \subset A \cdot eAe \subset Ae
  \]
  and so \(AL\) is a left ideal of \(A\) contained in \(Ae\), but
  \(Ae\) is irreducible, so \(AL = Ae\). Furthermore, since \(e\) acts
  as the identity on \(eAe\), then \(eL = L\) and \[
    eAe = eAL = eA(eL) = eAe \cdot L \subset L \implies eAe = L
  \]
  and so \(eAe\) has only \((0)\) and itself as left ideals. Thus,
  \(eAe\) is a division ring. \\

  Let us assume \(eAe\) is a division ring. Since \(A\) is semisimple,
  if \(Ae\) is not irreducible, then \[
    Ae = L_1 \oplus L_2
  \]
  where \(L_1, L_2\) are proper submodules. Then, this gives us the
  decomposition \(e = e_1 + e_2\) for idempotents \(e_1 \in L_1, e_2 \in
  L_2\). Note that
  \begin{align*}
    e & = e^2 \\
      &= (e_1 + e_2)(e_1 + e_2) \\
    & = e_1^2 + e_2 e_1 + e_1 e_2 + e_2^2 \\
    \implies 0 & = e_1^2 - e_1 + e_2 e_1 + e_1 e_2 + e_2^2 -e_2
  \end{align*}
  However, \(e_1 e_2 + e_2^2 - e_2 \in L_2\) and \(-e_2 e_1 - e_1^2 + e_1 \in L_1\) tells us that
  \(e_1 e_2 + e_2^2-e_2= - e_2 e_1 - e_1^2 + e_1= 0\) since \(L_1 \intersect L_2 =
  0\). Furthermore, \(e_1, e_2 \in eAe\) since \(e_1, e_2 \in Ae
  \implies e_1 e = e_1\) and \(e_2 e = e_2\), as well as
  \begin{align*}
    e e_2 & = (e_1 + e_2)e_2 \\
              & = e_1 e_2 + e_2^2 \\
              & = e_2
  \end{align*}
  and similarly \(e e_1 = e_1\). Thus, \(e_1 = e e_1 e \in eAe\) and
  \(e_2 = e e_2 e \in eAe\). Finally,
  \begin{align*}
    e_2^2 + e_1 e_2 & = e_2 \\
                    & = e_2(e_1+e_2) \\
                    & = e_2 e_1 +  e_2^2 \\
    \implies e_1 e_2 & = e_2 e_1
  \end{align*}
  But \(e_1 e_2 \in L_2\) and \(e_2 e_1 \in L_1\) and thus \(e_1 e_2 =
  0\), which tells us that \(eAe\) has zero divisors. Thus, \(eAe\) is
  not a division ring.

  % consider the
  % projection map \(\pi_1\) given by \[
  %   \begin{tikzcd}
  %     L_1 \oplus L_2 \ar[r,"\pi_1"] \ar[rd, twoheadrightarrow]& L_1 \oplus L_2 \\
  %     & L_1 \ar[u, hookrightarrow]
  %   \end{tikzcd}
  % \]
  % and similarly for \(\pi_2\). Then, \(0 \neq \pi_1, \pi_2
  % \in \End_A(Ae)\) but \(\pi_1 \pi_2 = 0\). 
\end{proof}
\section{The Structure of Semisimple (Left) Artinian Rings}
\begin{defn}
  A left artinian ring \(A\) is called \de{simple} if its only
  two-sided ideals
  are \(0\) and \(A\).
\end{defn}
\begin{defn}
  A set of idempotents \(\{e_1, \ldots, e_n\}\) of a ring \(A\) is a
  \de{complete set of orthogonal idempotents} if
  \begin{itemize}
  \item Each \(e_i\) is an idempotent,
  \item \(e_1 + \cdots + e_n = 1\)
  \item and \(e_i e_j = 0\) if \(i \neq j\).
  \end{itemize}
\end{defn}
Note that, for an artinian ring \(A\), we have \[
  {}_A A = L_1 \oplus \cdots \oplus L_n
\]
for \(L_i\) irreducible \(A\)-modules, also seen as minimal left
ideals. Thus, \[
  1 = e_1 + \cdots + e_n, \ \ e_i \in L_i
\]
with \(e_j e_i \in L_i\). Thus, \(e_j^2 = e_j\) and \(e_i e_j = 0\) if
\(i \neq j\), so \(\{e_1, \ldots, e_n\}\) is a complete set of
orthogonal idempotents for \(A\). \\

We now seek to prove the Artin-Wedderburn theorem, that will tell us
that a semisimple left artinian ring is isomorphic to a direct sum of
matrix rings over division rings.
\begin{lem}\label{wedderburn-lem-1}
  Let \(L,L'\) be minimal left ideals. Then, \(L \isom L'\) as left
  \(A\)-modules if and only if there is an \(a' \in L'\) such that
  \(L' = La'\).
\end{lem}
\begin{proof}
  For (\(\implies\)), let \(\phi \from L \isomto L'\). Then, for \(x
  \in L = (e)\), \[
    \phi(x) = \phi(xe) = x \phi(e) = xa' \ \text{ where } a' = \phi(e)
  \]
  Then, \(L' = La'\) as required. \\

  For (\(\impliedby\)), assume there is an \(a' \in L'\) such that
  \(L' = La'\). Define \(\phi \from L \to L'\) by \(\phi(x) =
  xa'\). Thus, \(\phi\) is surjective. Since \(L\) is irreducible,
  \(\ker \phi = \{0\}\) and so \(\phi\) is injective. Thus, \(\phi\)
  is an isomorphism.
\end{proof}
% \begin{lem}\label{wedderburn-lem-2}
%   Let \(L,L'\) be minimal left ideals in \(A\). Then, \[
%     L \isom L' \iff LL' = L'
%   \]
% \end{lem}
% \begin{proof}
%   For (\(\implies\)), we note by the lemma above that there exists an
%   \(a' \in L'\) such that \(La' = L'\). Thus, \(LL' \supset L'\) and moreover,
%   \(LL' \subset L'\) by standard ring theory. \\

%   For (\(\impliedby\)), assume \(LL' = L'\). Then, there is an \(a'
%   \in L'\) such that \(La' \neq 0\). However, \(L'\) is irreducible,
%   so \(La' = L'\) and so we are done by the above lemma.
% \end{proof}
\begin{thm}[Weddurburn]
  Let \(A\) be a semisimple artinian ring. Let \(L\) be a minimal left
  ideal. The sum of all minimal left ideals \(\{L' \ideal A
  \st L' \isom L\}\), say \(B_L\), is a simple subring and a
  \(2\)-sided ideal in \(A\). Also, \[
    A = \bigoplus_{i=1}^r B_{L_i}
  \]
  where \(L_1, \ldots, L_r\) are representatives of distinct
  isomorphism classes of left ideals.
\end{thm}
\begin{proof}
  Let \[
    A = L_1 \oplus \cdots \oplus L_r
  \]
  where the \(L_i\) are minimal left ideals of \(A\). There is an
  equivalence relation on \(\{L_1, \ldots, L_r\}\) given by \[
    L_i \sim L_j \iff L_i \isom L_j
  \]
  Thus, if \(L_1, \ldots, L_s\) is an equivalence class, then for
  \(i,j \leq s\), \(L_i L_j = L_j\). Thus, let us define \[
    B_{L_1} := L_{1} \oplus \cdots \oplus L_{s} \subset A 
  \]
  Such a subset is closed under multiplication and is a 2-sided ideal
  since \(A\). Let us do this 
  analogously for every equivalence class to get \(B_1, B_2, \ldots,
  B_t\). Then, \[
    i \neq j \implies B_i B_j = 0
  \]
  and also, \[
    B_1 \oplus B_2 \oplus \cdots \oplus B_t = A
  \]
  For a given \(B_i\), we have \[
    B_i = L_{i,1} \oplus L_{i,2} \oplus \cdots \oplus L_{i,s_i}
  \]
  where \(L_{i,j} = L_\ell\) for some \(\ell\). We note \[
    1_i := e_{i,1} + \cdots + e_{i,s_i}
  \]
  is the multiplicative identity in \(B_i\). Thus, \(B_i\) is a
  subring of \(A\). Also, \(B_i\) is a left ideal, so it is
  artinian.\\
  
  \todo{It is not clear to me where we show \(B_i\) is a 2-sided
    ideal.}
  
  It still remains to show that \(B_i\) is simple subring and a
  2-sided ideal. Without loss of generality, consider
  \begin{align*}
    B_1 & = L_1 \oplus \cdots \oplus L_r \\
    1_B & = e_1 + \cdots + e_r
  \end{align*}
  and let \(0 \neq D \ideal B_1\), that is, let \(D\) be a nontrivial
  2-sided ideal of \(B_i\). Then, \(D\) is a left-ideal and thus
  contains a minimal left ideal \(L\) of \(A\). Thus, \[
    L \subset D \subset B_1 \implies L \isom L_1
  \]
  and so we conclude that \(L \isom L_1\) since \(L \subset
  B_1\). However, we also know that \[ 
    Lx \subset D, \ \ x \in A
  \]
  since \(D\) is a right ideal, too. Thus, by \ref{wedderburn-lem-1},
  \(\{Lx \st x \in A\}\) gives all minimal left ideals of \(A\) isomorphic to
  \(L\). Therefore, it must be that \(D = B_1\), so \(B_1\) contains
  no non-trivial two-sided ideals and is thus a simple subring.
\end{proof}
\begin{thm}[Wedderburn]
  Let \(A\) be a simple artinian ring. Then, there exists a unique
  division ring and a unique positive integer \(n\) such that \[
    A \isom M_n(D)
  \]
  Conversely, if \(D\) is a division ring, then \(M_n(D)\) is a simple
  artinian ring. 
\end{thm}
Thus, combining our results, we arrive at the landmark theorem.
\begin{thm}[Artin-Wedderburn Theorem]
  Let \(A\) be a semisimple left artinian ring. Then, up to reordering,
  there is a unique decomposition \[
    A \isom M_{n_1}(D_1) \times M_{n_2}(D_2) \times \cdots \times M_{n_r}(D_r)
  \]
  where each \(D_i\) is a division ring and each \(n_i\) is a positive
  integer.
\end{thm}
\section{The Double Centralizer Property}
In order to finish our proofs, however, we will appeal to some other
facts that make use of the following language.
\begin{defn}
  Let \(A\) be a ring and \(M\) be a left \(A\)-module. Furthermore,
  let \(A_L \subset \End_A(M)\) be the endomorphisms given by applying
  elements of \(A\) to \(M\). If \(A_L = \End_A(M)\), we sa that the
  pair \((A,M)\) has the \de{double centralizer property}. 
\end{defn}
\begin{rmk}
  More generall, this notion can be extended to any ring \(R\) with
  subring \(S\). We say \((R,S)\) has the double centralizer property
  if \(C_R(C_R(S)) = S\).
\end{rmk}
Thus, Wedderburn's theorem above can be reprhased as follows
\begin{thm}[Wedderburn]
  Let \(A\) be a simple Artinian ring. Then, for some
  minimal left ideal \(M \ideal A\), \((A,M)\) has the double
  centralizer property.
\end{thm}
\begin{lem}
  Let \(A\) be a ring. Them \((A,{}_A A)\) has the double centralizer
  property. 
\end{lem}
\begin{lem}
  Let \(V = M^{\bigoplus k}\) for a left \(A\)-module \(M\), \(k \in
  \N\). If \((A,V)\) has the double centralizer property, then
  \((A,M)\) also has the double centralizer property.
\end{lem}
\begin{lem}
  Let \(M = Ae\) for \(e \in A\) an idempotent. Then, \[
    \End_A(M) \isom eAe
  \]
  via isomorphism \(f \from \End_A(M) \to eAe\) and \(md = m \cdot f(d)\) for all \(m \in M, d
  \in \End_A(M)\). 
\end{lem}
\begin{bibdiv}
  \begin{biblist}
    \bib{cr}{book}{
      author={Curtis, Curtis W.}
      author={Reiner, Irving}
      title={Representation Theory of Finite Groups and Associative
        Algebras}
      year={1962}
    }
    \bib{jacobson}{book}{
      author={Jacobson, Nathan}
      title={Basic Algebra II}
      year={1989}
      note={2nd edition}
    }
  \end{biblist}
\end{bibdiv}

\end{document}